Appendix S1: Supporting Tables and Figures

### 2022-04-27 ms/SI.Rmd
Update SI1 as per copy-editor changes

Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of RPD, SES of FD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*. Significant *I* > 0 indicates autocorrelation.

@@

Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix). Higher (less negative) log likelihood indicates better fit. Full environmental models included the effects of temperature, precipitation, precipitation seasonality, and area. Full reproductive models included the effects of % apomictic taxa, precipitation, precipitation seasonality, and area. PD, phylogenetic diversity; RPD, relative phylogenetic diversity; FD, functional diversity; RFD, relative functional diversity; SES, standard effect size.

@@

Model fit as measured with conditional Akaike information criterion (cAIC). Environmental models included the effects of temperature, precipitation, precipitation seasonality, and area. Reproductive models included the effects of % apomictic taxa, precipitation, precipitation seasonality, and area. Lower cAIC indicates better fit.

@@

# Distribution of redundancy by grain size
c <- ggplot(redundancy_by_res, aes(x = redundancy, y = res)) +
  geom_density_ridges(scale = 1.5) + 
  scale_y_discrete(expand = c(0, 0), labels = km_lab) + # will generally have to set the `expand` option
  scale_x_continuous(expand = c(0, 0)) +   # for both axes to remove unneeded padding
  coord_cartesian(clip = "off") + # to avoid clipping of the very top of the top ridgeline
  labs(
    x = "Redundancy",
    y = "Resolution"
  )

@@

Effect of grain size on sampling redundancy. (A) Percent of grid cells with redundancy > 0.5 by grain size (side length of square grid cells set to 10 km, 20 km, 30 km, or 40 km). (B) Percent of grid cells with redundancy = 0 by grain size. (C) Distribution of redundancy values by grain size.

@@

a <- plot_env("A", "Area", "lat_area", "km^2")
b <- plot_env("B", "% apomictic taxa", "percent_apo", labels = function(x) scales::percent(x, accuracy = 1))
c <- plot_env("C", "Precipitation", "precip", "mm")
d <- plot_env("D", "Precip. season.", "precip_season" , "")
e <- plot_env("E", "Temperature", "temp", "°C", function(x) x/10)
f <- plot_env("F", "Temp. season.", "temp_season") +
  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))

@@

Reproductive mode and environmental data in 20 km × 20 km grid cells across Japan. (A) rolling mean area (km^2^) in 20 km latitudinal bands. (B) % apomictic taxa. (C) annual precipitation. (D) percipitation seasonality (coefficient of variation). (E) mean annual temperature. (F) temperature seasonality (standard deviation × 100). Data excluded from models not shown in this and subsequent plots unless otherwise mentioned.

@@

Time tree of native Japanese ferns and lycophytes, excluding hybrids. Points at nodes indicate bootstrap support. Scalebar 100 my. Tree visualized with the `r pack("ggtree")` R package [@Yu2017].

@@

Observed number of specimens (A) and sampling redundancy (B) per 20 km × 20 km grid cell in the ferns of Japan.

@@

(A) Raw taxonomic diversity, (B) phylogenetic diversity, and (C) functional diversity of the ferns of Japan plotted by latitude. For (B) and (C), raw branch lengths were transformed to relative values by dividing each branchlength by the sum of all branch lengths. Trendlines fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].

@@

```{r endemism-restricted, fig.width = 5, fig.height = 5}
ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_endemic_spatial, aes(fill = endem_type), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  scale_fill_manual(
    values = canape_cols,
    na.value = "transparent") +
  annotation_scale(location = "tl", width_hint = 0.2) +
  map_theme() +
  theme(legend.title = element_blank())
```

@@

Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleoendemism), restricted data set including only taxa endemic to Japan. 'neo' indicates neodendemic areas (those with an overabundance of short branches); 'paleo' indicates paleoendemic areas (those with an overabundance of long branches); 'mixed' indicates significantly endemic sites that have neither an overabundance of short nor long branches; 'super' indicates highly significantly endemic sites that have neither an overabundance of short nor long branches. For details, see Materials and Methods.

@@

Selection of *k* for bioregions. (A) taxonomic bioregions. (B) phylogenetic bioregions. Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal).

@@

Bioregion dendrograms. Tips are sites (names not shown). Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between sites (see Materials and Methods). Color shows bioregion; bioregions not consisting of more than `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category. (A) taxonomic bioregions. (B) phylogenetic bioregions. Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].
\clearpage

@@

percent_apo_genus %>%
  filter(total_apo > 1) %>%
  group_by(genus) %>%
  summarize(
    n = mean(n),
    total_apo = mean(total_apo),
    percent_apo = mean(percent_apo),
    .groups = "drop"
  ) %>%
  mutate(
    genus =  paste0("*", genus, "*"),
    genus = fct_reorder(genus, percent_apo),
    label = glue::glue("{total_apo}/{n}")) %>%
  ggplot(aes(x = percent_apo, y = genus)) + 
  geom_col() +
  geom_text(aes(label = label), hjust = -.3) +
  scale_x_continuous(labels = percent, limits = c(0, .85)) +
  labs(x = "% apomictic taxa") +
  theme(
    axis.text.y = element_markdown(),
    axis.title.y = element_blank()
  )
```

@@

Rate of apomixis in genera of native, non-hybrid Japanese ferns with > 1 apomictic taxon. Numbers next to each bar show number of apomictic taxa out of total richness per genus.

@@

Relationship between (A) phylogenetic diversity (PD) and (B) relative PD and % apomictic taxa, models inferred using phylogeny with branch lengths in units of expected genetic change (untransformed ML tree). Ribbon shows 95% confidence interval of model. Line fit with the focal predictor variable while averaging over other predictors. For RPD, the standard effect size (SES) was calculated by comparing raw values to a null distribution of 999 random values, and significance (*P*-value) calculated using a two-tailed test (see Materials and Methods).

@@

Scatterplots of grid cell membership in (A) taxonomic or (B) phylogenetic bioregions arranged by mean annual temperature and annual precipitation. Points shown with partial transparency; darker areas indicate overlapping points. Bioregions not consisting of more than `r english2(bioregion_cutoff)` grid cells each are lumped into the "Other" category.

@@

Maps showing overlap of grid cells with significantly high biodiversity for the ferns of Japan and protected areas. Grid cells with black outlines indicate significantly high biodiversity. Biodiversity metrics include (A) taxon richness, (B) phylogenetic diversity (PD), (C) functional diversity (FD), and (D) phylogenetic endemism (PE). Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid cells in the top 5% considered highly diverse (see Materials and Methods).

>>>>

Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of RPD, SES of FD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.
Significant *I* > 0 indicates autocorrelation.
SES, standard effect size; PD, phylogenetic diversity; RPD, relative PD; FD, functional diversity; RFD, relative FD.

@@

Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix) for spatial models describing the effect of environmental or biological variables on the biodiversity of the ferns of Japan.
Higher (less negative) log likelihood indicates better fit.
Full environmental models included the effects of temperature, precipitation, precipitation seasonality, and area.
Full reproductive models included the effects of % apomictic taxa, precipitation, precipitation seasonality, and area.
PD, phylogenetic diversity; RPD, relative PD; FD, functional diversity; RFD, relative FD; SES, standard effect size.

@@

Model fit as measured with conditional Akaike information criterion (cAIC) for spatial models describing the effect of environmental or biological variables on the biodiversity of the ferns of Japan.
Environmental models included the effects of temperature, precipitation, precipitation seasonality, and area.
Reproductive models included the effects of % apomictic taxa, precipitation, precipitation seasonality, and area.
Lower cAIC indicates better fit.
PD, phylogenetic diversity; RPD, relative PD; SES, standard effect size.

@@

# Distribution of redundancy by grain size
c <- ggplot(redundancy_by_res, aes(x = redundancy, y = res)) +
  geom_density_ridges(scale = 1.5) + 
  scale_y_discrete(expand = c(0, 0), labels = km_lab) + # will generally have to set the `expand` option
  scale_x_continuous(expand = c(0, 0)) +   # for both axes to remove unneeded padding
  coord_cartesian(clip = "off") + # to avoid clipping of the very top of the top ridgeline
  labs(
    x = "Redundancy",
    y = "Resolution (km)"
  )

@@

Effect of grain size on sampling redundancy of the ferns of Japan. (A) Percentage of grid cells with redundancy >0.5 by grain size (side length of square grid cells set to 10 km, 20 km, 30 km, or 40 km). (B) Percentage of grid cells with redundancy = 0 by grain size. (C) Distribution of redundancy values by grain size.

@@

a <- plot_env("A", "Area", "lat_area", "km^2")
b <- plot_env("B", "% Apomictic taxa", "percent_apo", labels = function(x) scales::percent(x, accuracy = 1))
c <- plot_env("C", "Precipitation", "precip", "mm")
d <- plot_env("D", "Precip. season.", "precip_season" , "")
e <- plot_env("E", "Temperature", "temp", "°C", function(x) x/10)
f <- plot_env("F", "Temp. season.", "temp_season") +
  annotation_scale(location = "tl", width_hint = 0.2, pad_y = unit(0.8, "cm"))

@@

Reproductive mode and environmental data in 20 km × 20 km grid cells for the ferns of Japan.
(A) Rolling mean area (square km) in 20 km latitudinal bands.
(B) Percentage apomictic taxa.
(C) Annual precipitation.
(D) Percipitation seasonality (coefficient of variation).
(E) Mean annual temperature.
(F) Temperature seasonality (standard deviation × 100).
Data excluded from models not shown in this and subsequent plots unless otherwise mentioned.

@@

Time tree of native Japanese ferns and lycophytes, excluding hybrids.
Points at nodes indicate bootstrap support.
Scalebar = 100 Myr.
Tree visualized with the `r pack("ggtree")` R package [@Yu2017].

@@

Observed number of specimens (A) and sampling redundancy (B) per 20 km × 20 km grid cell for ferns of Japan.

@@

(A) Raw taxonomic diversity, (B) phylogenetic diversity (PD), and (C) functional diversity (FD) of the ferns of Japan plotted by latitude.
For (B) and (C), raw branch lengths were transformed to relative values by dividing each branchlength by the sum of all branch lengths.
Trendlines fit with general additive model using `r func("geom_smooth")` function in the `r pack("ggplot2")` R package [@Wickham2016].

@@

```{r endemism-restricted, fig.width = 5, fig.height = 5}
ggplot() +
  geom_sf(data = japan, fill = "white", color = "transparent", lwd = coast_lwd) +
  geom_sf(data = biodiv_ferns_endemic_spatial, aes(fill = endem_type), color = "transparent") +
  geom_sf(data = japan, fill = "transparent", color = coast_col, lwd = coast_lwd) +
  scale_fill_manual(
    values = canape_cols,
    breaks = names(canape_cols)[names(canape_cols) != "not significant"],
    name = "Endemism type") +
  annotation_scale(location = "tl", width_hint = 0.2) +
  map_theme()
```

@@

Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleoendemism), restricted data set including only taxa endemic to Japan.
neo, neodendemic areas (those with an overabundance of short branches); paleo, paleoendemic areas (those with an overabundance of long branches); mixed, significantly endemic sites that have neither an overabundance of short nor long branches; super, highly significantly endemic sites that have neither an overabundance of short nor long branches.
Grid cells without significant endemism in grey.
For details, see Materials and Methods.

@@

Selection of *k* for bioregions in the ferns of Japan. (A) Taxonomic bioregions. (B) Phylogenetic bioregions. Color indicates optimal *k* selected selected by the `r func("optimal_phyloregion")` function in the R package `r pack("phyloregion")` (black, non-optimal; red, optimal).

@@

Bioregion dendrograms for the ferns of Japan.
Tips are grid cells (names not shown).
Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid cells (see Materials and Methods).
Color shows bioregion; bioregions not consisting of more than `r english2(bioregion_cutoff)` grid cells each are lumped into the "Other" category.
(A) Taxonomic bioregions.
(B) Phylogenetic bioregions.
Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].

@@

percent_apo_genus %>%
  filter(total_apo > 1) %>%
  group_by(genus) %>%
  summarize(
    n = mean(n),
    total_apo = mean(total_apo),
    percent_apo = mean(percent_apo),
    .groups = "drop"
  ) %>%
  mutate(
    genus =  paste0("*", genus, "*"),
    genus = fct_reorder(genus, percent_apo),
    label = glue::glue("{total_apo}/{n}")) %>%
  ggplot(aes(x = percent_apo, y = genus)) + 
  geom_col() +
  geom_text(aes(label = label), hjust = -.3) +
  scale_x_continuous(labels = percent, limits = c(0, .85)) +
  labs(x = "% Apomictic taxa") +
  theme(
    axis.text.y = element_markdown(),
    axis.title.y = element_blank()
  )
```

@@

Rate of apomixis in genera of native, non-hybrid Japanese ferns with >1 apomictic taxon.
Numbers next to each bar show number of apomictic taxa out of total richness per genus.

@@

Relationship between (A) phylogenetic diversity (PD) and (B) relative PD (RPD) and % apomictic taxa in the ferns of Japan, models inferred using phylogeny with branch lengths in units of expected genetic change (untransformed ML tree).
Ribbon shows 95% confidence interval of model.
Line fit with the focal predictor variable while averaging over other predictors.
For RPD, the standard effect size (SES) was calculated by comparing raw values to a null distribution of 999 random values, and significance (*P*-value) calculated using a two-tailed test (see Materials and Methods).

@@

Scatterplots of grid cell membership in (A) taxonomic or (B) phylogenetic bioregions of the ferns of Japan arranged by mean annual temperature and annual precipitation.
Points shown with partial transparency; darker areas indicate overlapping points.
Bioregions not consisting of more than `r english2(bioregion_cutoff)` grid cells each are lumped into the "Other" category.

@@

Maps showing overlap of grid cells with significantly high biodiversity and protected areas for the ferns of Japan .
Grid cells with black outlines indicate significantly high biodiversity.
Biodiversity metrics include (A) taxon richness, (B) phylogenetic diversity (PD), (C) functional diversity (FD), and (D) phylogenetic endemism (PE).
Significance of PD, FD, and PE assessed by a one-tailed test comparing observed values to a null distribution of 999 random values; for richness, grid cells in the top 5% considered highly diverse (see Materials and Methods).
